Supplemental Materials: \emph{Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours}

### 2015-03-13 Publications/2015-NeuroImage/04-methods.tex
one more iteration to second round

\subsection{Susceptibility-derived distortions}
\label{sec:distortions}
\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
In order to quickly probe the diffusion process within the brain in many different orientations,
  these sequences minimize the bandwidth of the \gls*{pe} direction.
This low bandwidth originates the artifactual displacement of locations presenting small deviations
  from the main magnetic field $B_0$.
The principal source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
  susceptibility at tissue interfaces.
Tissue/air boundaries show large steps of susceptibility and this artifact is
  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
  bone of both hemispheres.
The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
  \citep{jezzard_correction_1995}:

  \begin{equation}
  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE},
  \label{eq:fieldmap}
  \end{equation}
%
where $\vec{r}$ is the true location in the undistorted domain as in \eqref{eq:transform},
  $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
  $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
  vector along \gls*{pe}.
Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
The first solution consisted on estimating $\Delta B_0$ with field mapping
  techniques \citep{andersson_modeling_2001}.
A second family of methods require an extra \gls*{epi} acquisition, but switching
  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
\cite{kybic_unwarping_2000} proposed a correction using nonlinear registration of the \emph{b0}
  image\footnote{The term \emph{b0} refers to one or more images within the
  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
   \emph{b}-value used as signal reference.} of the \gls*{dmri} versus a \gls*{t2} image.
Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
  three-dimensional problem.
Therefore, a typical protocol for the extraction of structural connectivity
  comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
  method of choice \citep{daducci_connectome_2012}.

\subsection{Our method and contributions}
\label{sec:our_method}
\emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
  surfaces that can be extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012},
  to then map this prior information into a target image.
However, \emph{bbregister} features an affine transform that is unable to compensate the
  distortions.
To avoid this problem, the active contours are dismissed in regions typically affected by
  warpings.
Additionally, \emph{bbregister} is robust in finding clear tissue boundaries.
In the case of \gls*{dmri} data, oftentimes the registration process is restricted to finding
  the pial surface, as it is the only clear boundary.

to the methods based on nonlinear registration of the
  to the \gls*{t2} image, since we also render the correction as a
  registration problem.
Conversely, the method does not require any of the previously mentioned extra acquisitions.
A fundamental antecedent to \emph{regseg} is found in .
They proposed a registration method to align surfaces extracted from the \gls*{t1}
  to \gls*{dmri} images using a linear transformation model using active contours
  with edges.
Similarly, we  warped data, in which the scale of the
  imaged structures is in the range of the image resolution or above.
Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
First, we use a nonlinear deformation model that enables the correction of the
  susceptibility-derived distortions \citep{jezzard_correction_1995}.
\cite{greve_accurate_2009} circumvented the nonlinearity by identifying typically
  distorted regions and removing them from the metric computation.
Second, our matching criteria does not necessarily need for clear intensity
  steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
  active contours without edges \citep{chan_active_2001}.

This choice is similar to \citep{guyader_combined_2011}, but \emph{regseg} differs in two
  features.
First, the regularization relies on the inherent smoothness of the deformation field,
  and on the anisotropic approach proposed by \cite{nagel_investigation_1986}.
Second, it accepts several contours to be mapped as in \citep{gorthi_active_2011}.
However, \emph{regseg} makes use of the explicit representation of surfaces, and
  does not implement multiphase level sets.
Our registration method is demonstrated here as a solution to correct for the susceptibility
  distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
The distortion correction and the precise segmentation of \gls*{dmri} data is crucial in
	applications such as connectome extraction \citep{craddock_imaging_2013}, since the
  definition of the nodes in the network is generally derived from a parcellation performed in
  an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.

@@

  \begin{equation}
  P(U \mid R,\Gamma_{l,m}) = \frac{P(R \mid U,\Gamma_{l,m})\, P(U)}{P(R)},
  \label{eq:bayes_rule}
  \end{equation}

  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
  between piecewise smooth regions $\Omega_i$, such that
  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
  contour between competing regions $\Omega_l$ and $\Omega_m$.
The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
  \citep{bishop_pattern_2006}, and aligns $\Gamma$ into $R$.

@@

  \begin{equation}
  P(R \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}

  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

 	\begin{align}
  P( R \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}

  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.

@@

  \begin{align*}
  P(U) = \underset{\vec{r}}{\prod}\, p(u(\vec{r})) &=
  \underset{\vec{r}}{\prod}\, p_0(u(\vec{r})) \, p_1(u(\vec{r})), \\
  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, \mathbf{A}^{-1}), \notag\\
  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, \mathbf{B}^{-1}),
  \end{align*}

  imposing that the distortion and its gradient have zero
  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
Since the anisotropy is generally aligned with the imaging axes, these will be simplified
  in the following for the sake of clarity:

  \begin{align}
    p_0(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( u(\vec{r}) \mid 0,
      (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}), \notag\\
    p_1(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{r}) \mid 0,
      (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1}).
  \label{eq:priors}
  \end{align}

@@

  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
  &= C + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} \right] \,d\vec{r},
  \label{eq:energy}
  \end{align}

  this expression is dual to the energy functional corresponding
  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
  with the corresponding anisotropic regularization term \citep{nagel_investigation_1986}.

@@

\paragraph*{Deformation model}\label{sec:deformation_model}
Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
  surface(s).
In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
  \citep{fischl_freesurfer_2012}.
The transform $\hat{U}$ \eqref{eq:transform} is supported by a dense deformation field
  $u(\vec{r})$, such that:

@@

Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$:

  \begin{equation}
  u(\vec{r}) = \sum_k \psi_k(\vec{r}) \vec{u}_k.
  \label{eq:intp_kernel}
  \end{equation}

@@

  \begin{equation}
    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
  \label{eq:transformation}
  \end{equation}

  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.

@@

  \begin{equation}
  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}

  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} ] \,d\vec{r}
  \Big\}.
  \label{eq:gradient_descent}
  \end{align}

@@

  \begin{align}
  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \left\{
   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\}
  = \underset{i}{\sum}
   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
  \label{eq:gradient_wshape}
  \end{align}

  in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
    being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
    them such that
    $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
    and $\bar{s}_i'$ is the speed vector projected on to the unit inward normal to the contour
    at $c_i'$ (described in \autoref{fig:method}).

@@

  \begin{equation}
  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{equation}

  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:

@@

\todo[inline]{Include here the final update equation using the FT, introducing the
step size $\tau$, and refer to the supplementals for intermediate steps}

\paragraph*{Settings, implementation details, and convergence}
\label{sec:conv_report}
The registration parameters (such as $\tau$, the B-Spline grid resolutions,
  target image smoothing, etc.) and implementation details (for instance the sparse matrix approach
  to fast interpolation) are discussed in the \suppl{}, section S1.
The actual choices of parameter settings are publicly distributed with the source code of the experiments.
Additionally, we release our software along with a tool for generating convergence reports to
  demonstrate the behavior of \emph{regseg} and help scientists configure their own experiments.
One sample report is found in the \suppl{}, section S1.3.

>>>>

\subsection{Susceptibility-derived distortions}
\label{sec:distortions}
\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
In order to quickly probe the diffusion process within the brain in many different orientations,
  these sequences minimize the bandwidth of the \gls*{pe} direction.
This low bandwidth originates the artifactual displacement of locations presenting small deviations
  from the main magnetic field $B_0$.
The principal source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
  susceptibility at tissue interfaces.
Tissue/air boundaries show large steps of susceptibility and this artifact is
  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
  bone of both hemispheres.
Distortion can be rendered as a registration problem between the undistorted $R$ (reference), 
  and distorted $M$ (moving) spaces:

  \begin{align}
  U_s\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
  \label{eq:transform}
  \end{align}
%
  where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
  its corresponding location in $M$, and $n$ the dimensionality of images.
Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
  to the reference domain.

The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
  \citep{jezzard_correction_1995}:

  \begin{equation}
  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE},
  \label{eq:fieldmap}
  \end{equation}
%
where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
  $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
  vector along \gls*{pe}.
Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
The first solution consisted on estimating $\Delta B_0$ with field mapping
  techniques \citep{andersson_modeling_2001}.
A second family of methods require an extra \gls*{epi} acquisition, but switching
  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
\cite{kybic_unwarping_2000} proposed a correction using nonlinear registration of the \emph{b0}
  image\footnote{The term \emph{b0} refers to one or more images within the
  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
   \emph{b}-value used as signal reference.} of the \gls*{dmri} versus a \gls*{t2} image.
Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
  three-dimensional problem.
Therefore, a typical protocol for the extraction of structural connectivity
  comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
  method of choice \citep{daducci_connectome_2012}.

\subsection{Our method and contributions}
\label{sec:our_method}
% \emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
%   surfaces that can be extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012},
%   to then map this prior information into a target image.
% However, \emph{bbregister} features an affine transform that is unable to compensate the
%   distortions.
% To avoid this problem, the active contours are dismissed in regions typically affected by
%   warpings.
% Additionally, \emph{bbregister} is robust in finding clear tissue boundaries.
% In the case of \gls*{dmri} data, oftentimes the registration process is restricted to finding
%   the pial surface, as it is the only clear boundary.
% 
% to the methods based on nonlinear registration of the
%   to the \gls*{t2} image, since we also render the correction as a
%   registration problem.
% Conversely, the method does not require any of the previously mentioned extra acquisitions.
% A fundamental antecedent to \emph{regseg} is found in .
% They proposed a registration method to align surfaces extracted from the \gls*{t1}
%   to \gls*{dmri} images using a linear transformation model using active contours
%   with edges.
% Similarly, we  warped data, in which the scale of the
%   imaged structures is in the range of the image resolution or above.
% Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
% First, we use a nonlinear deformation model that enables the correction of the
%   susceptibility-derived distortions \citep{jezzard_correction_1995}.
% \cite{greve_accurate_2009} circumvented the nonlinearity by identifying typically
%   distorted regions and removing them from the metric computation.
% Second, our matching criteria does not necessarily need for clear intensity
%   steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
%   active contours without edges \citep{chan_active_2001}.
% This choice is similar to \citep{guyader_combined_2011}, but \emph{regseg} differs in two
%   features.
% First, the regularization relies on the inherent smoothness of the deformation field,
%   and on the anisotropic approach proposed by \cite{nagel_investigation_1986}.
% Second, it accepts several contours to be mapped as in \citep{gorthi_active_2011}.
% However, \emph{regseg} makes use of the explicit representation of surfaces, and
%   does not implement multiphase level sets.
% Our registration method is demonstrated here as a solution to correct for the susceptibility
%   distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
% The distortion correction and the precise segmentation of \gls*{dmri} data is crucial in
% 	applications such as connectome extraction \citep{craddock_imaging_2013}, since the
%   definition of the nodes in the network is generally derived from a parcellation performed in
%   an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.

@@

  \begin{equation}
  P(U \mid R,\Gamma_{l,m}) = \frac{P(R \mid U,\Gamma_{l,m})\, P(U)}{P(R)},
  \label{eq:bayes_rule}
  \end{equation}
%
  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
  between piecewise smooth regions $\Omega_i$, such that
  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
  contour between competing regions $\Omega_l$ and $\Omega_m$.
The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
  \citep{bishop_pattern_2006}, and aligns $\Gamma$ into $R$.

@@

  \begin{equation}
  P(R \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}
%
  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

 	\begin{align}
  P( R \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
%
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.

@@

  \begin{align*}
  P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \\
  p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
  p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
  \end{align*}
%
  imposing that the distortion and its gradient have zero
  mean and variance governed by the matrices $\mathbf{A} = (a_{ij})$ and $\mathbf{B} = (b_{ij})$.
Since the anisotropy is generally aligned with the imaging axes, $\mathbf{A}$ and $\mathbf{B}$
  can be simplified to diagonal matrices.
For convenience, we will set $\alpha_i = a_{ii}^2$ and $\beta_i = b_{ii}^2$, what can be reduced
  using the Hadamard square root notation\footnote{The Hadamard power of a matrix or a vector
  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$} as
  $\mathbf{A}=\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n$ and
  $\mathbf{B}=\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n$.

@@

  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
  &= C + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r},
  \label{eq:energy}
  \end{align}
%
  this expression is dual to the energy functional corresponding
  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
  with the corresponding anisotropic regularization term \citep{nagel_investigation_1986}.

@@

\paragraph*{Deformation model}\label{sec:deformation_model}
Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
  surface(s).
In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
  \citep{fischl_freesurfer_2012}.
The transform $\hat{U}$ \eqref{eq:transform} is supported by a dense deformation field
  $\vec{u}$, such that:

@@

Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $\vec{u}$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$:

  \begin{equation}
  \vec{u} = \sum_k \psi_k(\vec{r}) \vec{u}_k.
  \label{eq:intp_kernel}
  \end{equation}

@@

  \begin{equation}
    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
  \label{eq:transformation}
  \end{equation}
%
  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.

@@

  \begin{equation}
  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:general_gradient_descent}
  \end{equation}
%
  with $t$ being an artificial time parameter of the contour
  evolution, and $\vec{u}_k$ the parameters supporting the estimate
  $\hat{U}$ of the transformation at the current time point.
Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
  \Big\}.
  \label{eq:gradient_descent}
  \end{align}

@@

  \begin{align}
  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \left\{
   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\}
  = \underset{i}{\sum}
   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
  \label{eq:gradient_wshape}
  \end{align}
%
  in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
    being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
    them such that
    $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
    and $\bar{s}_i'$ is the speed vector projected on to the unit inward normal to the contour
    at $c_i'$ (described in \autoref{fig:method}).

@@

  \begin{equation}
  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{equation}
%
  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:

@@

Finally, we introduce a step size parameter $\delta$ to discretize the artificial time parameter
  in the iterative optimization.
Then, applying the Fourier transform as detailed in the \suppl{}, section S1, the
  update equation is derived from \eqref{eq:gradient_final}:

  \begin{align}
  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\vec{u}_k^t / \delta - \sum_i \vec{g}_{i,k}\}}%
                  {\mathcal{F}\{(1/\delta+\boldsymbol{\alpha})I-\boldsymbol{\beta}\Delta\}} \right\}
  \label{eq:update_equation}
  \end{align}

\paragraph*{Settings, implementation details, and convergence}
\label{sec:conv_report}
The registration parameters (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$,
  the B-Spline grid resolutions, target image smoothing, etc.)
  and other implementation details (for instance the sparse matrix approach
  to fast interpolation) are discussed in the \suppl{}, section S1.
The actual choices of parameter settings are publicly distributed with the source code of the experiments.
Additionally, we release our software along with a tool for generating convergence reports to
  demonstrate the behavior of \emph{regseg} and help scientists configure their own experiments.
One sample report is found in the \suppl{}, section S1.3.

### 2015-03-13 Publications/2015-NeuroImage/06-discussion.tex
one more iteration to second round

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
In section \nameref{sec:methods_background}, we framed our method within the current
  state of art and defined the contributions in our approach.
We chose active contours without edges as we were interested in the homogeneous
  regions of diffusion properties within the brain.
Iteratively, the active contours are updated following the instant shape gradients
  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
The approach is valid for image processing applications in which a precise segmentation is
  available (from an atlas or within subject anatomical data), and match any or all
  the following circumstances:
  i) reference and target images do not present appropriate contrasts to
  	perform volume-based registration reliably;
  ii) the low resolution of the target volume produces strong partial volume effects
  	that preclude voxel-wise segmentation algorithms and hinder volume-based
  	registration; and,
  iii) the estimation of the unknown warping is explicitly needed and the
    deformation field is smooth enough to be represented with a sufficient number
  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
  	object tracking in time series, etc.).
To support our claim that images can be simultaneously segmented and registered to a
  reference coordinate system, we conducted evaluation experiments on digital
  simulated phantoms and real datasets.

>>>>

% -*- root: 00-main.tex -*-
\section*{Discussion}
\label{sec:discussion}
% In section \nameref{sec:methods_background}, we framed our method within the current
%   state of art and defined the contributions in our approach.
% We chose active contours without edges as we were interested in the homogeneous
%   regions of diffusion properties within the brain.
% Iteratively, the active contours are updated following the instant shape gradients
%   (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
% The approach is valid for image processing applications in which a precise segmentation is
%   available (from an atlas or within subject anatomical data), and match any or all
%   the following circumstances:
%   i) reference and target images do not present appropriate contrasts to
%   	perform volume-based registration reliably;
%   ii) the low resolution of the target volume produces strong partial volume effects
%   	that preclude voxel-wise segmentation algorithms and hinder volume-based
%   	registration; and,
%   iii) the estimation of the unknown warping is explicitly needed and the
%     deformation field is smooth enough to be represented with a sufficient number
%   	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
%   	object tracking in time series, etc.).
% To support our claim that images can be simultaneously segmented and registered to a
%   reference coordinate system, we conducted evaluation experiments on digital
%   simulated phantoms and real datasets.
